### 2022-04-09 paper-swahler-jaiio2022.tex
Se agrega descripción de datos

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage[outdir=./]{epstopdf}
\usepackage{graphicx}
\usepackage[spanish,es-noshorthands]{babel}
%\usepackage{Tikz}
\usepackage{pspicture}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%
\begin{document}
%
\input{titulo}
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
\input{Abstract}
\end{abstract}
%
%
\section{Introducción}
\input{Introduccion}

\section{Análisis de Redes Sociales (SNA)}
\input{sna}

\section{Marco de Trabajo - Ministerio Público Fiscal del Chubut}
\input{marco-trabajo}

\section{Descripción general de los Datos}
\input{descripcion-datos}

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\subsubsection{Acknowledgements} Please place your acknowledgments at
the end of the paper, preceded by an unnumbered run-in heading (i.e.
3rd-level heading).

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{bibliografia}
%
\renewcommand{\refname}{Referencias Bibliográficas}
\begin{thebibliography}{8}
\bibitem{ref_article1}
Finckenauer JO. Problems of definition: What is organized crime? Trends in Organized Crime. 2005; 8(3):63–83. \url{https://doi.org/10.1007/s12117-005-1038-4}
\bibitem{ref_article2}
Krishna Raj P.M.Ankith MohanK.G. Srinivasa. Practical Social Network Analysis with Python. Springer, 2018. ISBN: 978-3-319-96746-2.
\bibitem{ref_article3}
Burcher, Morgan. Social Network Analysis and Law Enforcement: Applications for Intelligence Analysis. Serie Crime Prevention and Security Management, 2020. ISBN 978-3-030-47770-7
\bibitem{ref_article4}
L.C. Freeman, The SAGE handbook of social network analysis, eds. J. Scott, P.J. Carrington, SAGE Publications Ltd., 2011.
\bibitem{ref_article5}
W.R. Harper, D.H. Harris, The application of link analysis to police intelligence. Hum. Factors 17(2), 157–164 (1975).
\bibitem{ref_article6}
P. Klerks, The network paradigm applied to criminal organisations: theoretical nitpicking or relevant doctrine for investigators? Recent developments in the Netherlands. Connections 24(3), 53–65 (1999)
\bibitem{ref_article7}
V. Krebs, Mapping networks of terrorist cells. Connections 24(3), 43–52 (2002)
\bibitem{ref_article8}
R.M. Medina, Social network analysis: a case study of the Islamist terrorist network. Secur. J. 27(1), 97–121 (2014)
\bibitem{ref_article9}
J. Qin, J.J. Xu, D. Hu, M. Sageman, H. Chen, Analyzing terrorist networks: a case study of the global jihad. Lect. Notes Comput. Sci. 3495, 287–304 (2005)
\bibitem{ref_article10}
E. Stollenwerk, T. Dörfler, J. Schibberges, Taking a new perspective: mapping the Al Qaeda network through the eyes of the UN Security Council. Terror. Political Violence 28(5), 950–970  (2016)
\bibitem{ref_article11}
M. Bouchard, J. Amirault, Advances in research on illicit networks. Glob. Crime 14(2–3), 119–122 (2013)
\bibitem{ref_article12}
D.A. Bright, C. Greenhill, M. Reynolds, A. Ritter, C. Morselli, The use of actor-level attributes and centrality measures to identify key actors: a case study of an Australian drug trafficking network. J. Contemp. Crim. Justice 31(3), 262–278 (2015a)
\bibitem{ref_article13}
L. Giommoni, A. Aziani, G. Berlusconi, How do illicit drugs move across countries? a network analysis of the heroin supply to Europe. J. Drug Issues 47(2), 217–240 (2016)
\bibitem{ref_article14}
C. Morselli, Hells Angels in springtime. Trends Org. Crime 12(2), 145–158 (2009)
\bibitem{ref_article15}
C. Morselli, Assessing vulnerable and strategic positions in a criminal network. J. Contemp. Crim. Justice 26(4), 382–392 (2010)
\bibitem{ref_article16}
A. Malm, G. Bichler, Networks of collaborating criminals: assessing the structural vulnerability of drug markets. J. Res. Crime Delinq. 48(2), 271–297 (2011)
\bibitem{ref_article17}
G. Bichler, A. Malm, Small arms, big guns: a dynamic model of illicit market opportunity. Glob. Crime 14(2–3), 261–286 (2013)
\bibitem{ref_article18}
A.F. Colladon, E. Remondi, Using social network analysis to prevent money laundering. Expert Syst. Appl. 67, 49–58 (2017)
\bibitem{ref_article19}
M.R.J. Soudijn, Using strangers for money: a discussion on money-launderers in organized crime. Trends Org. Crime 17(3), 199–217 (2014)
\bibitem{ref_article20}
M. Lauchs, R. Keast, N. Yousefpour, Corrupt police networks: uncovering hidden relationship patterns, functions and roles. Polic. Soc. 21(1), 110–127 (2011)
\bibitem{ref_article21}
J.M. McGloin, Policy and intervention considerations of a network analysis of street gangs. Criminol. Public Policy 4(3), 607–635 (2005)
\bibitem{ref_article22}
G. Bichler, A. Malm, J. Enriquez, Magnetic facilities: identifying key juvenile convergence places with social network analysis. Crime Delinq. 60(7), 971–998 (2014)
\bibitem{ref_article23}
D. Décary-Hétu, Information exchange paths in IRC hacking chat rooms, in Crime and networks, ed. by C. Morselli, (Routledge, New York, 2014)
\bibitem{ref_article24}
D. Décary-Hétu, B. Dupont, The social network of hackers. Glob. Crime 13(3), 160–175 (2012)
\bibitem{ref_article25}
D. Décary-Hétu, B. Dupont, Reputation in a dark network of online criminals. Glob. Crime 14(2–3), 175–196 (2013)
\bibitem{ref_article26}
M1. Xu, Jennifer, and Hsinchun Chen. "Criminal network analysis and visualization." Communications of the ACM 48.6 (2005): 100-107.
\bibitem{ref_article27}
M2. Feng, Mingchen, et al. "Big data analytics and mining for effective visualization and trends forecasting of crime data." IEEE Access 7 (2019): 106111-106123.
\bibitem{ref_article28}
M3. Mathew, Ammu, et al. "Criminal Networks Mining and Visualization for Crime Investigation." Caniya and P, Mufeed and Raj, Asha, Criminal Networks Mining and Visualization for Crime Investigation (July 8, 2021) (2021).
\bibitem{ref_article29}
M4. Chen, Hsinchun, et al. "Visualization in law enforcement." CHI'05 extended abstracts on Human factors in computing systems. 2005.
\bibitem{ref_article30}
\url{https://www.mpfchubut.gov.ar/}
\bibitem{ref_article31}
L. González, G. Rua. Sistemas Judiciales: Una perspectiva integral sobre la administración de justicia - Análisi Criminal, Publicación anual del CEJA e INECIP. N°23. Año 2019.

\end{thebibliography}
\end{document}