### 2022-04-26 titulo.tex

\title{Análisis Inteligente de Datos y Visualización aplicadas a la Investigación Criminal\\
Intelligent Data Analysis and Visualization applied to Criminal Investigation
\thanks{Universidad Nacional del Sur - Bahía Blanca - Argentina. Universidad Nacional de la Patagonia San Juan Bosco - Chubut - Argentina.}}
%
\titlerunning{Análisis y Visualización de Redes de Datos aplicada a la Justicia

Analysis and Visualization of Data Networks applied to Justice}
%

% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sebastián P. WAHLER\inst{1} \and
	Diego C. MARTÍNEZ\inst{1} \and
	Martín L. LARREA\inst{1}}
%
\authorrunning{S. Wahler, D. Martínez, M. Larrea}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Departamento de Ciencias e Ingeniería de la Computación, Universidad Nacional del Sur, Av. Alem 1253, B8000CPB Bahía Blanca, ARGENTINA.
	\email{spwahler@ing.unp.edu.ar, dcm@cs.uns.edu.ar, mll@cs.uns.edu.ar}\\
	\url{https://cs.uns.edu.ar/}}
%