### 2022-05-09 titulo.tex

\title{Análisis y Visualización de Redes Sociales para identificar Bandas Delictivas\\
Analysis and Visualization of Social Networks to identify Criminal Gangs
\thanks{Universidad Nacional de la Patagonia San Juan Bosco - Chubut - Argentina.\\Universidad Nacional del Sur - Bahía Blanca - Argentina.}}
%
\titlerunning{Análisis y Visualización de Bandas Delictivas}

%
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Sebastián P. WAHLER\inst{1} \inst{2} \and Diego C. MARTÍNEZ\inst{3} \and Martín L. LARREA\inst{3}}
%
\authorrunning{S. Wahler, D. Martínez, M. Larrea}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
	\inst{}Departamento de Informática, Facultad de Ingeniería, Universidad Nacional de la Patagonia San Juan Bosco, Mitre 655, U9100, Trelew, ARGENTINA.\\
	\url{http://www.ing.unp.edu.ar/dpto-informatica.html}
	\and
	\inst{}Departamento de Informática, Procuración General, Ministerio Público Fiscal, Poder Judicial de la Provincia del Chubut, Belgrano 521, U9102, Rawson, ARGENTINA.\\
	\url{https://www.mpfchubut.gov.ar}
	\and
	\inst{}Departamento de Ciencias e Ingeniería de la Computación, Universidad Nacional del Sur, Av. Alem 1253, B8000CPB Bahía Blanca, ARGENTINA.\\
	\url{https://cs.uns.edu.ar/} \\
	\email{(e-mail: spwahler@ing.unp.edu.ar, dcm@cs.uns.edu.ar, mll@cs.uns.edu.ar)}\\	
}

%